Automatic Hadith Graphs using Arabic Morphological Analysis

### 2011-08-04 trunk/tex/hadith.ijcnlp.tex
changed the biography clustering code (now calculates the center also)

{\bf Sarf.~~~}
Sarf~\footnote{Sarf (name modified for blind review) }
is 
an {\em application-specific efficient
morphological analyzer} based on
non-deterministic finite state Automata 
driven by an application-specific controller customized 
by the user. 
%Each machine in Sarf takes one letter at a time as input
%and represents a valid morphological analysis.
The application-specific controller prunes false positives
early in the run by making a decision at every input letter
to control the machines.
Sarf builds on and extends the lexicon of 
Buckwalter~\shortcite{Buckwalter:02} 
with proper and location names extracted from different online 
sources~\footnote{\href{http://alasmaa.net/}{http://alasmaa.net/ }, 
\href{http://ar.wikipedia.org/}{http://ar.wikipedia.org/}}
%as well as biblical sources~\footnote{Genesis 4:17-23; 5:1-32; 9:28-10:32; 11:10-32; 25:1-4, 12-18; 36:1-37:2; Exodus 6:14-25; Ruth 4:18-22; 1 Samuel 14:49-51; 1 Chronicles 1:1-9:44; 14:3-7; 24:1; 25:1-27:22; Nehemiah 12:8-26; Matthew 1:1-16; Luke 3:23-38}.
as well as biblical sources.
%~\footnote{Genesis 4:17-23; 5:1-32; 9:28-10:32; 11:10-32; 25:1-4, 12-18; 36:1-37:2; Exodus 6:14-25; Ruth 4:18-22; 1 Samuel 14:49-51; 1 Chronicles 1:1-9:44; 14:3-7; 24:1; 25:1-27:22; Nehemiah 12:8-26; Matthew 1:1-16; Luke 3:23-38}.

@@

In this paper we make the following contributions. 
\begin{itemize}
\item We automate the segmentation of hadith books into separate narrations and extract chains of narrators into a partial order relation graph using a knowledge based approach with morphological analysis.
\item We introduce a specialized distance metric for Arabic names used in Arabic literature. 
\item We provide a working and efficient implementation of
our work. 
\end{itemize}

@@

@@

@@

@@

\transfalse
\novocalize
\begin{figure*}[tb]
\center{
\resizebox{1.8\columnwidth}{!}
%{ \input{iTreeMohamada}}
{\includegraphics{iTreeMohamada}}
\caption{Output of iTree fails to detect \RL{m_hmdA}.}
\label{f:mohamada}
}
\end{figure*}
\transtrue

@@

To evaluate our approach, we ran NaGEMA against three books of 
hadith~\cite{IbnHanbal,AlTousi,AlKulayni} and compared the
results to iTree. 
The results are detailed in 
Tables~\ref{t:hadithresults} and \ref{t:nopunctlearnresults}. 
The numbers we present in the tables are the result of a 
manual verification of the correctness of the outputs of 
10\% of the narrations in the books under consideration.

@@

First, we compare the recall of the segmentation 
in NaGEMA to that of hadith acceptance rates in iTree. 
Although the task of segmentation is more complex than that of
hadith recognition/acceptance, we still notice that iTree has on the average 35.7\% recall rate compared to 99.9\% for 
NaGEMA. 
However, we notice that the difference in these numbers depended highly on the books under consideration. 
For instance,
for the book of musnad, both approaches correctly detected  practically all narrations under consideration; 
however iTree suffers a lot in 
the books of both Kafi and Istibsar, resulting in a recall of less than 5\%. 
This is due to the less structured nature of the narrations of those books.
For example, iTree's grammar is defined such that a hadith starts by 
an Ikhbar word (\noVocRL{AdaT al-A_hbAar}) such as
\noVocRL{qAl}, \noVocRL{.hadda_t}, etc...; 
however, the Kafi and Istibsar books followed a flexible style to express the narration 
and most narrations started directly by a narrator name without a preceding Ikhbar word. 
This is a strong evidence of the superiority and tolerance of NaGEMA 
and its threshold based approach
compared to the strict grammar approach used in iTree. 
Clearly, NaGEMA does not fail on such cases, since what it detects as a hadith is based 
on a concentration of names.
Similarly, iTree had a lower sanad boundary precision (62.4\% on the average).
This is mainly due to the grammar that ends a sanad only when a predefined prophet phrase is met.
Prophet phrases are phrases that point to the prophet such as 
\noVocRL{rasowl al-lah} and 
\noVocRL{.sal_A `alayh al-lah wasalam}.

@@

\subsection{On Unorganized hadith books}

(report results on such hadith books)

@@

(show detailed statistics and compare against that of eNarrator)

@@

>>>>

{\bf Sarf.~~~}
Sarf~\footnote{Sarf (name modified for blind review) }
is 
an {\em application-specific efficient
morphological analyzer} based on
non-deterministic finite state Automata 
driven by an application-specific controller customized 
by the user. 
%Each machine in Sarf takes one letter at a time as input
%and represents a valid morphological analysis.
The application-specific controller prunes false positives
early in the run by making a decision at every input letter
to control the machines.
Sarf builds on and extends the lexicon of 
Buckwalter~\shortcite{Buckwalter:02} 
with proper and location names extracted from different online 
sources~\footnote{\href{http://alasmaa.net/}{http://alasmaa.net/ }, 
\href{http://ar.wikipedia.org/}{http://ar.wikipedia.org/}}
as well as biblical sources~\footnote{Genesis 4:17-23; 5:1-32; 9:28-10:32; 11:10-32; 25:1-4, 12-18; 36:1-37:2; Exodus 6:14-25; Ruth 4:18-22; 1 Samuel 14:49-51; 1 Chronicles 1:1-9:44; 14:3-7; 24:1; 25:1-27:22; Nehemiah 12:8-26; Matthew 1:1-16; Luke 3:23-38}.

@@

In this paper we make the following contributions. 
\begin{itemize}
\item We automate the segmentation of hadith books into separate narrations and extract chains of narrators 
%into a partial order relation graph 
using a knowledge based approach with morphological analysis.
\item We optimize our analysis through utilizing punctuation and learning previously unknown names.
\item We build from different chains of narrators a partial order relation graph based on merging common narrators.
\item We introduce a specialized distance metric for Arabic names used in Arabic literature. 
\item We highly outperform the currently available narrator extraction tool. 
\item We provide a working and efficient implementation of our work. 
\end{itemize}

@@

(Figure~\ref{f:chain} is confusing since it misses some information uses other symbols like 
those of $\theta$ and is redundant with what follows)

@@

(explain better and merged information present about them in "Results" section)

@@

\subsection{Speedup Optimizations for Large Datasets}

\subsubsection{MultiKey Hashing}

(explain what are the keys used for each value and why suitable)

@@

%\transfalse
%\novocalize
%\begin{figure*}[tb]
%\center{
%\resizebox{1.8\columnwidth}{!}
%%{ \input{iTreeMohamada}}
%{\includegraphics{iTreeMohamada}}
%\caption{Output of iTree fails to detect \RL{m_hmdA}.}
%\label{f:mohamada}
%}
%\end{figure*}
%\transtrue

@@

To evaluate our approach, we ran NaGEMA against three books of 
hadith~\cite{IbnHanbal,AlTousi,AlKulayni} and compared the
results to iTree. 
The results are detailed in 
Tables~\ref{t:hadithresults} and \ref{t:nopunctlearnresults}. 
The numbers we present in the tables are the result of a 
manual verification of the correctness of %the outputs of 
10\% of the narrations in the books under consideration.

(discuss annotation and verification process and mention smg about inter-annotator agreement)

@@

First, we compare the recall of the segmentation 
in NaGEMA to that of hadith acceptance rates in iTree. 
Although the task of segmentation is more complex than that of
hadith recognition/acceptance, we still notice that iTree has on the average 35.7\% recall rate compared to 99.9\% for 
NaGEMA. 
However, we notice that the difference in these numbers depended highly on the books under consideration. 
For instance,
for the book of musnad, both approaches correctly detected  practically all narrations under consideration; 
however iTree suffers a lot in 
the books of both Kafi and Istibsar, resulting in a recall of less than 5\%. 
This is due to the less structured nature of the narrations of those books.
For example, iTree's grammar is defined such that a hadith starts by 
an Ikhbar word (\noVocRL{AdaT al-A_hbAar}) such as
\noVocRL{qAl} ``said'', \noVocRL{.hadda_t} ``narrated'', etc...; 
however, the Kafi and Istibsar books followed a flexible style to express the narration 
and most narrations started directly by a narrator name without a preceding Ikhbar word. 
This is a strong evidence of the superiority and tolerance of NaGEMA 
and its threshold based approach
compared to the strict grammar approach used in iTree. 
Clearly, NaGEMA does not fail on such cases, since what it detects as a hadith is based 
on a concentration of names.
Similarly, iTree had a lower sanad boundary precision (62.4\% on the average).
This is mainly due to the grammar that ends a sanad only when a predefined prophet phrase is met.
Prophet phrases are phrases that point to the prophet such as 
\noVocRL{rasowl al-lah} and 
\noVocRL{.sal_A `alayh al-lah wasalam}.

@@

%\subsection{On Unorganized hadith books}
%
%(report results on such hadith books)

@@

(show detailed statistics and compare against that of eNarrator which uses a similarity string comparison appraoch)

@@

\section{Significance \& Extensions -- Named Entity Extraction in Other Contexts}

\subsection{Genealogical Lists in the Bible}

(explain what will be equivalent of NMC and NRC and how it will work)

(Here is a sophisticated example:
``This is the book of Adam’s history. In the day of God’s creating Adam he made him in the likeness of God. Male and female he created them. After that he blessed them and called their name Man in the day of their being created. And Adam lived on for a hundred and thirty years. Then he became father to a son in his likeness, in his image, and called his name Seth. And the days of Adam after his fathering Seth came to be eight hundred years. Meanwhile he became father to sons and daughters. So all the days of Adam that he lived amounted to nine hundred and thirty years and he died. And Seth lived on for a hundred and five years. Then he became father to E′nosh. And after his fathering E′nosh Seth continued to live eight hundred and seven years. Meanwhile he became father to sons and daughters. So all the days of Seth amounted to nine hundred and twelve years and he died. And E′nosh lived on for ninety years. Then he became father to Ke′nan. And after his fathering Ke′nan E′nosh continued to live eight hundred and fifteen years. Meanwhile he became father to sons and daughters. So all the days of E′nosh amounted to nine hundred and five years and he died...'' (Genesis 5:1-11)

Here is a simple example:

``Adam, Seth, E′nosh,  Ke′nan, Ma‧hal′a‧lel, Ja′red,  E′noch, Me‧thu′se‧lah, La′mech,  Noah, Shem, Ham and Ja′pheth. The sons of Ja′pheth were Go′mer and Ma′gog and Ma′da‧i and Ja′van and Tu′bal and Me′shech and Ti′ras. And the sons of Go′mer were Ash′ke‧naz and Ri′phath and To‧gar′mah. And the sons of Ja′van were E‧li′shah and Tar′shish, Kit′tim and Ro′da‧nim. The sons of Ham were Cush and Miz′ra‧im, Put and Ca′naan.'' (1 Chronicles 1:1-8)
)

In many cases, such lists cover overlapping sections and when they do, they sometimes exhibit some seeming contradictions 
due to nicknames and other reasons. Detecting those conflicts is a task that has been performed by countless 
number of Bible students, theologians and historians. 
Their results have been published in articles and books; however, almost all of them are in written format 
and rarely displayed the genealogical graphs to make it easier for the reader to pinpoint those conflicts and visualize them.

In addition, being able to create a comprehensive genealogical graph from all biblical accounts would greatly enhance its study,
since readers would be able to better understand a character's genealogical context and life span. Such comprehensive graphs can be tagged
with time durations between a father and his giving birth of his sons as is common in many biblical accounts...

\subsection{Organizing Quoted Email Replies/Forwards}

(we can use the "FROM:" and punctuation to extract the sender and the body of the email and organize a 
received email into individual chains of replies/forwards. Also the Inbox can be merged into a POR of contacts with whom a user communicates and the history of their communication, as in the case of the hadith...)

\subsection{Place Names in Travel Itineraries}

(check if it can be applicable)

(Here is an example from wikipedia ($http://wikitravel.org/en/Silk_Road$):

-Northern Silk Route \\
Dunhuang to Hami \\
Hami to Turfan \\ 
Turfan to Korla \\
Korla to Kashgar

-Middle Silk Route \\ 
Dunhuang to Cherchen \\
Cherchen to Korla \\
Korla to Kashgar

-Southern (Jade) Silk Route \\
Dunhuang to Cherchen \\
Cherchen to Khotan \\
Khotan to Yarkand \\
Yarkand to Kashgar \\
)

In many cases, these itineraries contain some common places and merging them in a POR is much more illustrative than having them in text format.
Also, when electronic maps are available, such textual itineraries can automatically be turned into emphasized routes on maps.

\subsection{Extending Arabic People Name Gazetteers}

(briefly explain how this specific application has general value through contributing to the general task of entity extraction [learned names])
